---
title: "Motor Trend Magazine - MPG"
author: "Allan Perk"
date: "21 September 2016"
output: html_document
---

```{r setup, include=FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## Executive Summary

This article's main point is to analyze fuel consumption in miles per gallon (MPG) based on trasnmission type, along with other potential impactors. 
To perform the anlaysis, we first conduct an exploratory data analysis, after which we investigat and perform the best regression. To assess the best choice, we use the *leaps* package to assess explanotory power and attempt to be parsimonious. We conclude by testing the data for normality using a *t.test* and *qqplot*.

## Data
The data used is the *mtcars* dataset available in R, which was extracted by the 1974 Motor Trend Magazine, which presents fuel consumption and 10 other aspects of automobile design and performance for a total of 32 automobiles. 

A data frame with 32 observations on 11 variables.

[, 1]	 mpg	 Miles/(US) gallon
[, 2]	 cyl	 Number of cylinders
[, 3]	 disp	 Displacement (cu.in.)
[, 4]	 hp	 Gross horsepower
[, 5]	 drat	 Rear axle ratio
[, 6]	 wt	 Weight (1000 lbs)
[, 7]	 qsec	 1/4 mile time
[, 8]	 vs	 V/S
[, 9]	 am	 Transmission (0 = automatic, 1 = manual)
[,10]	 gear	 Number of forward gears
[,11]	 carb	 Number of carburetors

```{r data, echo = FALSE, warning = FALSE}
library(pacman)
p_load(ggplot2, GGally, gridExtra, leaps, corrplot, pastecs, knitr)
data(mtcars)
str(mtcars)
head(mtcars)
```

## Exploratory Analysis

Using the pastecs package for descriptive statistics, we can have a good idea on whether the data follows a normal distribution. With the median approximately equal to the mean, the skew.2SE less than one, and kurtosis approximately zero we can just about conclude the data follows a normal distribution. Additionally, taking the p.value of the normtest at 0.12288, we cannot reject the nul of the distribution being close to normal.


```{r pressure, echo=FALSE, warning = FALSE}
stat.desc(mtcars[1], basic=FALSE, norm=TRUE)

c <- cor(mtcars)
cp <- corrplot(c, type="lower", diag = TRUE)

grid.arrange(st, cp, nrow = 1, ncol = 2)

g = ggplot(mtcars, aes(mpg))
h = g+ geom_histogram(aes(y=..density..),fill = "light blue", stat = "bin", binwidth= 2) + geom_density() 
fg = h + facet_wrap(~am)

grid.arrange(h, fg, nrow = 2, ncol = 1)           
b = ggplot(mtcars, aes(y=mpg, x=am)) + geom_boxplot()
```

## Regression Analysis

```{r, echo = FALSE}
#Test for normality
t.test(mtcars$mpg)
t.test(mpg~am, mtcars)

           
#Regression Models
fitaut <- lm(mpg~am, mtcars)
fitall <- lm(mpg~., mtcars)
lower = fitaut
upper = fitall
fitstp <- step(fitaut, scope = list(lower=lower, upper=upper), direction = "forward")  

#Alternative using leaps
leaps = regsubsets(mpg~., mtcars)
plot(leaps, scale="adjr2")
#Residual Analysis
par(mfrow=c(2,2))


#QQplot            
ggplot(mtcars, aes(sample = mpg)) + 
  stat_qq(geom = "point", aes(colour= am)) + ggtitle("QQPLOT Miles per gallon")

```
